UniScatter: Unifying\\Backscatter Communications, Symbiotic Radio and Reconfigurable Intelligent Surface

### 2022-08-16 main.tex
:pencil: Update signal model

% glossaries-extra
\setabbreviationstyle[acronym]{long-short}
\newacronym{af}{AF}{Amplify-and-Forward}
\newacronym{ambc}{AmBC}{Ambient Backscatter Communications}
\newacronym{ap}{AP}{Access Point}
\newacronym{awgn}{AWGN}{Additive White Gaussian Noise}
\newacronym{bcd}{BCD}{Block Coordinate Descent}
\newacronym{bc}{BackCom}{Backscatter Communications}
\newacronym{bibo}{BIBO}{Binary-Input Binary-Output}
\newacronym{bpcu}{\si{bpcu}}{bits per channel use}
\newacronym{bpsphz}{\si{bps/Hz}}{bits per second per Hertz}
\newacronym{cp}{CP}{Canonical Polyadic}
\newacronym{cr}{CR}{Cognitive Radio}
\newacronym{cscg}{CSCG}{Circularly Symmetric Complex Gaussian}
\newacronym{csi}{CSI}{Channel State Information}
\newacronym{df}{DF}{Decode-and-Forward}
\newacronym{dmc}{DMC}{Discrete Memoryless Channel}
\newacronym{dmtc}{DMTC}{Discrete Memoryless Thresholding Channel}
\newacronym{dmtmac}{DMTMAC}{Discrete Memoryless Thresholding Multiple Access Channel}
\newacronym{dtmac}{DTMAC}{Discrete Thresholding Multiple Access Channel}
\newacronym{dp}{DP}{Dynamic Programming}
\newacronym{fdma}{FDMA}{Frequency-Division Multiple Access}
\newacronym{iid}{i.i.d.}{independent and identically distributed}
\newacronym{ioe}{IoE}{Internet of Everything}
\newacronym{iot}{IoT}{Internet of Things}
\newacronym{kkt}{KKT}{Karush-Kuhn-Tucker}
\newacronym{m2m}{M2M}{Machine to Machine}
\newacronym{mac}{MAC}{Multiple Access Channel}
\newacronym{mc}{MC}{Multiplication Coding}
\newacronym{miso}{MISO}{Multiple-Input Single-Output}
\newacronym{mimo}{MIMO}{Multiple-Input Multiple-Output}
\newacronym{ml}{ML}{Maximum-Likelihood}
\newacronym{mrt}{MRT}{Maximum Ratio Transmission}
\newacronym{noma}{NOMA}{Non-Orthogonal Multiple Access}
\newacronym{ofdm}{OFDM}{Orthogonal Frequency-Division Multiplexing}
\newacronym{pdf}{PDF}{Probability Density Function}
\newacronym{pgd}{PGD}{Projected Gradient Descent}
\newacronym{psk}{PSK}{Phase Shift Keying}
\newacronym{qam}{QAM}{Quadrature Amplitude Modulation}
\newacronym{qos}{QoS}{Quality of Service}
\newacronym{rf}{RF}{Radio-Frequency}
\newacronym{rfid}{RFID}{Radio-Frequency Identification}
\newacronym{ris}{RIS}{Reconfigurable Intelligent Surface}
\newacronym{sc}{SC}{Superposition Coding}
\newacronym{sic}{SIC}{Successive Interference Cancellation}
\newacronym{simo}{SIMO}{Single-Input Multiple-Output}
\newacronym{sinr}{SINR}{Signal-to-Interference-plus-Noise Ratio}
\newacronym{smawk}{SMAWK}{Shor-Moran-Aggarwal-Wilber-Klawe}
\newacronym{snr}{SNR}{Signal-to-Noise Ratio}
\newacronym{sr}{SR}{Symbiotic Radio}
\newacronym{swipt}{SWIPT}{Simultaneous Wireless Information and Power Transfer}
\newacronym{tdma}{TDMA}{Time-Division Multiple Access}
\newacronym{ue}{UE}{user}
\newacronym{wit}{WIT}{Wireless Information Transfer}
\newacronym{wpcn}{WPCN}{Wireless Powered Communication Network}
\newacronym{wpt}{WPT}{Wireless Power Transfer}

@@

	\emph{Notations:}
	Italic, bold lower-case, and bold upper-case letters denote scalars, vectors and matrices, respectively.
	$\boldsymbol{0}$ and $\boldsymbol{1}$ denote zero and one array of appropriate size, respectively.
	$\mathbb{R}_+^{x \times y}$ and $\mathbb{C}^{x \times y}$ denote the real nonnegative and complex spaces of dimension $x \times y$, respectively.
	$j$ denotes the imaginary unit.
	% $\arg(\cdot)$, $\mathrm{rank}(\cdot)$, $\mathrm{tr}(\cdot)$, $\mathrm{diag}(\cdot)$ and $\mathrm{diag}^{-1}(\cdot)$ denote the argument, rank, trace, a square matrix with input vector on main diagonal, and a vector retrieving main diagonal of input square matrix, respectively.
	$\mathrm{diag}(\cdot)$ returns a square matrix with the input vector on its main diagonal and zeros elsewhere.
	$(\cdot)^*$, $(\cdot)^T$, $(\cdot)^H$, $\lvert{\cdot}\rvert$, and $\lVert{\cdot}\rVert$ denote the conjugate, transpose, conjugate transpose, absolute value, and Euclidean norm operators, respectively.
	$(\cdot)^{(r)}$ and $(\cdot)^{\star}$ denote the $r$-th iterated and terminal solutions, respectively.
	The distribution of a \gls{cscg} random variable with zero mean and variance $\sigma^2$ is denoted by $\mathcal{CN}(0,\sigma^2)$, and $\sim$ means ``distributed as''.
\end{section}

@@

	\begin{subsection}{BackCom/SR: Backscatter Modulation}
		\gls{bc} and \gls{sr} commonly employ antenna-based scatterers as passive information sources.
		By \emph{frequently switching} between the available states, each backscatter node encodes self message over scattered signal and introduces backscatter uncertainty at the receiver.
		For $M$-ary \gls{qam}, reflection coefficient $\Gamma_m$ maps to the corresponding \emph{complex constellation point $c_m$} by
		% \cite{Thomas2012a}
		\begin{equation}
			\Gamma_m = \alpha \frac{c_m}{\max_{m'} \lvert c_{m'} \rvert},
			\label{eq:backscatter_modulation}
		\end{equation}
		where $0 \le \alpha \le 1$ is the amplitude scattering ratio at the direction of interest.
	\end{subsection}

	\begin{subsection}{RIS: Channel Reconfiguration}
		\gls{ris} typically involves a planar antenna array or metasurface of numerous metamaterial units as a passive channel reconfigurator.
		By \emph{adaptively choosing} the reflection pattern, each reflecting element alters the phase of scattered signal for constructive/destructive superposition at the receiver.
		For each unit with $M$ candidate states, reflection coefficient $\Gamma_m$ maps to the corresponding \emph{phase shift $\theta_m$} by
		% \cite{Wu2018}
		\begin{equation}
			\Gamma_m = \beta_m \exp(j \theta_m),
			\label{eq:passive_beamforming}
		\end{equation}
		where $0 \le \beta_m \le 1$ is the overall amplitude scattering ratio of state $m$.
	\end{subsection}

@@

\begin{section}{UniScatter}
	% Using shared spectrum, energy and infrastructures, UniScatter is a novel passive communication/assistance protocol that generalizes \gls{bc}, \gls{sr} and \gls{ris} from the perspective of distribution.
	% Using shared spectrum, energy and infrastructures, UniScatter is a novel passive communication/assistance protocol that generalizes \gls{bc}, \gls{sr} and \gls{ris} via \emph{physical} and \emph{probability} distributions.
	Using shared spectrum, energy and infrastructures, UniScatter is a novel passive communication/assistance protocol that generalizes \gls{bc}, \gls{sr} and \gls{ris} in a \emph{physically} and \emph{probabilistically} distributed manner.
	In particular, passive UniScatter nodes located at different places ride over an active legacy transmission in a flexible and mutualistic manner, while practical UniScatter receiver cooperatively decodes both primary and backscatter links under the benefits of backscatter modulation and passive beamforming.
	% TODO compare BC, SR and IRS
	\begin{subsection}{UniScatter Nodes}
		Based on link priority and \gls{csi}, UniScatter nodes adapt the \emph{state input probability distribution} of antenna-based scatterers to unify backscatter modulation and passive beamforming.
		The scattering sources of \gls{bc}/\gls{sr} and reflecting elements of \gls{ris} can be viewed as its extreme cases, where the input distribution boils down to equiprobable and degenerate, respectively.
		\begin{figure}[!t]
			\centering
			\subfloat[Input Distribution]{
				\resizebox{0.8\columnwidth}{!}{
					\input{assets/illustration/input_distribution.tex}
				}
				\label{fi:input_distribution}
			}
			\\
			\subfloat[Time Block]{
				\resizebox{0.8\columnwidth}{!}{
					\input{assets/illustration/time_block.tex}
				}
				\label{fi:time_block}
			}
			\caption{
				Input probability distribution and time block structure of \gls{bc}, \gls{sr}, \gls{ris}, and UniScatter.
				$T_s$ and $T_c$ respectively denote the primary and backscatter symbol period.
				Within channel coherence time, UniScatter nodes semi-randomly select reflection state for each backscatter symbol block under the guidance of input probability distribution.
			}
			\label{fi:uniscatter_distribution}
		\end{figure}
		As illustrated in Fig. \ref{fi:uniscatter_distribution}, instead of always using fully random or deterministic reflection pattern over time, each UniScatter node can \emph{flexibly adjust} its input distribution to balance information encoding and channel reconfiguration.
		That is, it semi-randomly select the reflection state per backscatter symbol block under the guidance of input probability $P(\Gamma_m)$ for state $m$.
		Such an adaptive backscatter channel coding coincides with \gls{ris} when primary link is absolutely prioritized, and outperforms the uncoded equiprobable inputs of \gls{bc}/\gls{sr} when backscatter link is absolutely prioritized.
		Besides, the physically distributed characteristics of UniScatter nodes and legacy users avoids the development optimization and allows uniformly good performance for both links.

@@

		Denote the \gls{ap}-user direct channel as $\boldsymbol{h}_{\mathrm{D}}^H \in \mathbb{C}^{1 \times Q}$, the \gls{ap}-node $k \in \mathcal{K} \triangleq \{1,\ldots,K\}$ forward channel as $\boldsymbol{h}_{\mathrm{F},k}^H \in \mathbb{C}^{1 \times Q}$, the node $k$-user backward channel as $h_{\mathrm{B},k}$, and the cascaded channel via tag $k$ as $\boldsymbol{h}_{\mathrm{C},k}^H \triangleq h_{\mathrm{B},k} \boldsymbol{h}_{\mathrm{F},k}^H \in \mathbb{C}^{1 \times Q}$.
		% Let $x_{\mathcal{K}} \triangleq \{x_k : k \in \mathcal{K}\}$ be the backscatter symbols of all UniScatters.
		Let $x_{\mathcal{K}} \triangleq (x_1,\ldots,x_K)$ be the backscatter symbol tuple of all UniScatter nodes.
		Consider the signal model during one backscatter block (i.e., $N$ primary blocks).
		% We assume the primary symbol follows standard \gls{cscg} distribution and the backscatter symbol of all tags employs $M$-\gls{qam}.
		Under perfect synchronization, the equivalent primary channel is a function of \emph{coded} backscatter symbols
		\begin{subequations}
			\label{eq:equivalent_channel}
			\begin{align}
				\boldsymbol{h}_{\mathrm{E}}^H(x_{\mathcal{K}})
				 & \triangleq \boldsymbol{h}_{\mathrm{D}}^H + \sum_{k \in \mathcal{K}} \alpha_k \boldsymbol{h}_{\mathrm{C},k}^H x_k   \\
				 & = \boldsymbol{h}_{\mathrm{D}}^H + \boldsymbol{x}^H \mathrm{diag}(\boldsymbol{\alpha}) \boldsymbol{H}_{\mathrm{C}},
			\end{align}
		\end{subequations}
		where $\alpha_k$ is the amplitude scattering ratio of UniScatter node $k$, $\boldsymbol{\alpha} \triangleq [\alpha_1,\ldots,\alpha_K]^T \in \mathbb{R}_+^{K \times 1}$, $x_k \in \mathcal{X} \triangleq \{c_1,\ldots,c_M\}$ is the coded backscatter symbol of node $k$, and $\boldsymbol{x} \triangleq [x_1,\ldots,x_K]^H \in \mathbb{C}^{K \times 1}$, $\boldsymbol{H}_{\mathrm{C}} \triangleq [\boldsymbol{h}_{\mathrm{C},1},\ldots,\boldsymbol{h}_{\mathrm{C},K}]^H \in \mathbb{C}^{K \times Q}$. The signal received by the user at primary block $n \in \mathcal{N} \triangleq \{1,\ldots,N\}$ is
		\begin{equation}
			y[n] = \boldsymbol{h}_{\mathrm{E}}^H(x_{\mathcal{K}}) \boldsymbol{w} s[n] + v[n],
		\end{equation}
		where $s \sim \mathcal{CN}(0,1)$ is the primary symbol, $v \sim \mathcal{CN}(0,\sigma_v^2)$ is the \gls{awgn}, and $\boldsymbol{w} \in \mathbb{C}^{Q \times 1}$ is the active beamforming vector with average power constraint $\lVert \boldsymbol{w} \rVert^2 \le P$.
		% The equivalent channel for primary link is subject to backscatter uncertainty
		% The equivalent primary channel is a function of backscatter symbols  as
		% \begin{equation}
		% 	% \boldsymbol{h}_{\mathrm{E}}^H(x_{\mathcal{K}}) \triangleq \boldsymbol{h}_{\mathrm{D}}^H + \sum_{k \in \mathcal{K}} \sqrt{\alpha_k} \boldsymbol{h}_{\mathrm{C},k}^H x_k.
		% 	\boldsymbol{h}_{\mathrm{E}}^H(\boldsymbol{x}) \triangleq \boldsymbol{h}_{\mathrm{D}}^H + \boldsymbol{x}^H \mathrm{diag}(\boldsymbol{\alpha}) \boldsymbol{H}_{\mathrm{C}}.
		% 	\label{eq:equivalent_channel}
		% \end{equation}

		% where $\alpha_k$ and $x_k$ denote respectively the harvest-backscatter efficiency and backscatter symbol of tag $k$, $s[n]$ and $w[n] \sim \mathcal{CN}(0,\sigma_v^2)$ denote respectively the primary symbol and \gls{awgn} at block $n$, and $\boldsymbol{w} \in \mathbb{C}^{Q \times 1}$ is the transmit precoder satisfying power constraint $\lVert \boldsymbol{w} \rVert^2 \le P$.
		% For the ease of notation, we define $x_{\mathcal{K}} \triangleq \{x_k : k \in \mathcal{K}\}$ as the tag input combination, and $\boldsymbol{y} \triangleq \left[y[1],\ldots,y[N]\right]^T \in \mathbb{C}^{N \times 1}$ as the received signal per backscatter symbol period.
		% For primary transmission, the equivalent channel is subject to backscatter modulation uncertainty as
		% \begin{equation}
		% 	\boldsymbol{h}_{\mathrm{E}}^H(x_{\mathcal{K}}) \triangleq \boldsymbol{h}_{\mathrm{D}}^H + \sum_{k \in \mathcal{K}} \sqrt{\alpha_k} \boldsymbol{h}_{\mathrm{C},k}^H x_k.
		% 	\label{eq:equivalent_channel}
		% \end{equation}

		% At a specific backscatter block, denote $m_k \in \mathcal{M}$ as the state index of UniScatter $k$, and let $m_{\mathcal{K}} \triangleq \{m_k: k \in \mathcal{K}\}$ collect the state indexes of all UniScatters.
		At a specific backscatter block, denote $m_k \in \mathcal{M}$ as the state index of node $k$, and let $m_{\mathcal{K}} \triangleq (m_1,\ldots,m_K)$ be the state index tuple of all nodes.
		% The received signal at primary block $n$ is subject to the variation of $s[n]$ and $v[n]$, and thus distributed as $y[n] \sim \mathcal{CN}(0,\sigma_v^2)$
		Conditioned on $m_{\mathcal{K}}$, the received signal at primary block $n$ is subject to the variation of $s[n]$ and $v[n]$, distributed as $y[n] \sim \mathcal{CN}(0,\sigma_{m_{\mathcal{K}}}^2)$ with

@@

			% \sigma_{m_{\mathcal{K}}}^2 = \Bigl\lvert \underbrace{\bigl(\boldsymbol{h}_{\mathrm{D}}^H + \sum_{k \in \mathcal{K}} \sqrt{\alpha_k} \boldsymbol{h}_{\mathrm{C},k}^H x_{m_k}\bigr)}_{\boldsymbol{h}_{\mathrm{E}}^H(x_{m_{\mathcal{K}}})} \boldsymbol{w} \Bigr\rvert^2 + \sigma_v^2,
			\sigma_{m_{\mathcal{K}}}^2 = \lvert \boldsymbol{h}_{\mathrm{E}}^H(x_{m_{\mathcal{K}}}) \boldsymbol{w} \rvert^2 + \sigma_v^2,

@@

		where $x_{m_k}$ and $x_{m_\mathcal{K}}$ are the symbol and symbol tuple associated with state $m_k$ and state tuple $m_{\mathcal{K}}$, respectively.%

@@

			$x_k$ and $x_{\mathcal{K}}$ are random variables, while $x_{m_k}$ and $x_{m_{\mathcal{K}}}$ are their instances indexed by $m_k$ and $m_{\mathcal{K}}$.

@@

		Also, denote the total received energy within backscatter block as $z=\sum_{n=1}^N \lvert y[n] \rvert^2$.
		% As the sum of $N$ \gls{iid} exponential variables, conditioned on $m_{\mathcal{K}}$, its \gls{pdf} follows Erlang distribution
		As the sum of $N$ \gls{iid} exponential variables, the \gls{pdf} of $z$ conditioned on $m_{\mathcal{K}}$ follows Erlang distribution

@@

			% f(z|\mathcal{H}_{m_{\mathcal{K}}}) = \frac{z^{N-1} \exp(\frac{-z}{\sigma_{m_{\mathcal{K}}}^2})}{\sigma_{m_{\mathcal{K}}}^{2N} (N-1)!},

@@

		% To accommodate backscatter characteristics and reduce decoding complexity, we consider a joint semi-coherent energy detection for all UniScatters based on disjoint decision regions over accumulated energy $z$.
		% !
		% To accommodate backscatter characteristics and reduce decoding complexity, we consider a joint semi-coherent detection for all UniScatter nodes over accumulated energy $z$.
		% % To reduce decoding complexity, we consider a joint semi-coherent energy detection for all UniScatters based on disjoint decision regions over accumulated energy $z$.
		% \begin{figure}[!t]
		% 	\centering
		% 	\resizebox{0.9\columnwidth}{!}{
		% 		\input{assets/illustration/energy_distribution.tex}
		% 	}
		% 	% \caption{\gls{pdf} of Average Received Power Conditioned on Different Input Hypothesis. Example \gls{ml}.}
		% 	\caption{
		% 		\gls{pdf} of total received energy per backscatter block, conditioned on different input hypothesis.
		% 		% This \gls{ml} decision consists of convex regions and is generally rate-suboptimal except for equiprobable inputs.}
		% 		Here, the convex \gls{ml} decision regions are generally rate-suboptimal except for equiprobable inputs.
		% 	}
		% 	\label{fi:energy_distribution}
		% \end{figure}
		% The idea is presented in Fig. \ref{fi:energy_distribution}.
		% !
		% As illustrated in Fig. \ref{fi:energy_distribution}, o
		Once disjoint energy decision regions are determined, we can construct a \gls{dtmac} and formulate the transition probability from input $x_{m_{\mathcal{K}}}$ to output $\hat{x}_{m_{\mathcal{K}}'}$ as
		% it essentially formulates a \gls{dtmac}, and the transition probability from input $x_{m_{\mathcal{K}}}$ to output $\hat{x}_{m_{\mathcal{K}}'}$ is

@@

			\label{eq:dtmac}

@@

		is the regularized incomplete Gamma function that substitutes the \gls{dtmac} transition probability \eqref{eq:dtmac}.

>>>>

% glossaries-extra
\setabbreviationstyle[acronym]{long-short}
\newacronym{af}{AF}{Amplify-and-Forward}
\newacronym{ambc}{AmBC}{Ambient Backscatter Communications}
\newacronym{ap}{AP}{Access Point}
\newacronym{awgn}{AWGN}{Additive White Gaussian Noise}
\newacronym{bcd}{BCD}{Block Coordinate Descent}
\newacronym{bc}{BackCom}{Backscatter Communications}
\newacronym{bibo}{BIBO}{Binary-Input Binary-Output}
\newacronym{bpcu}{\si{bpcu}}{bits per channel use}
\newacronym{bpsphz}{\si{bps/Hz}}{bits per second per Hertz}
\newacronym{cp}{CP}{Canonical Polyadic}
\newacronym{cr}{CR}{Cognitive Radio}
\newacronym{cscg}{CSCG}{Circularly Symmetric Complex Gaussian}
\newacronym{csi}{CSI}{Channel State Information}
\newacronym{df}{DF}{Decode-and-Forward}
\newacronym{dmc}{DMC}{Discrete Memoryless Channel}
\newacronym{dmtc}{DMTC}{Discrete Memoryless Thresholding Channel}
\newacronym{dmtmac}{DMTMAC}{Discrete Memoryless Thresholding Multiple Access Channel}
% \newacronym{dtmac}{DTMAC}{Discrete Thresholding Multiple Access Channel}
\newacronym{dp}{DP}{Dynamic Programming}
\newacronym{fdma}{FDMA}{Frequency-Division Multiple Access}
\newacronym{iid}{i.i.d.}{independent and identically distributed}
\newacronym{ioe}{IoE}{Internet of Everything}
\newacronym{iot}{IoT}{Internet of Things}
\newacronym{kkt}{KKT}{Karush-Kuhn-Tucker}
\newacronym{m2m}{M2M}{Machine to Machine}
\newacronym{mac}{MAC}{Multiple Access Channel}
\newacronym{mc}{MC}{Multiplication Coding}
\newacronym{miso}{MISO}{Multiple-Input Single-Output}
\newacronym{mimo}{MIMO}{Multiple-Input Multiple-Output}
\newacronym{ml}{ML}{Maximum-Likelihood}
\newacronym{mrt}{MRT}{Maximum Ratio Transmission}
\newacronym{noma}{NOMA}{Non-Orthogonal Multiple Access}
\newacronym{ofdm}{OFDM}{Orthogonal Frequency-Division Multiplexing}
\newacronym{pdf}{PDF}{Probability Density Function}
\newacronym{pgd}{PGD}{Projected Gradient Descent}
\newacronym{psk}{PSK}{Phase Shift Keying}
\newacronym{qam}{QAM}{Quadrature Amplitude Modulation}
\newacronym{qos}{QoS}{Quality of Service}
\newacronym{rf}{RF}{Radio-Frequency}
\newacronym{rfid}{RFID}{Radio-Frequency Identification}
\newacronym{ris}{RIS}{Reconfigurable Intelligent Surface}
\newacronym{sc}{SC}{Superposition Coding}
\newacronym{sic}{SIC}{Successive Interference Cancellation}
\newacronym{simo}{SIMO}{Single-Input Multiple-Output}
\newacronym{sinr}{SINR}{Signal-to-Interference-plus-Noise Ratio}
\newacronym{smawk}{SMAWK}{Shor-Moran-Aggarwal-Wilber-Klawe}
\newacronym{snr}{SNR}{Signal-to-Noise Ratio}
\newacronym{sr}{SR}{Symbiotic Radio}
\newacronym{swipt}{SWIPT}{Simultaneous Wireless Information and Power Transfer}
\newacronym{tdma}{TDMA}{Time-Division Multiple Access}
\newacronym{ue}{UE}{user}
\newacronym{wit}{WIT}{Wireless Information Transfer}
\newacronym{wpcn}{WPCN}{Wireless Powered Communication Network}
\newacronym{wpt}{WPT}{Wireless Power Transfer}

@@

	\emph{Notations:}
	Italic, bold lower-case, and bold upper-case letters denote scalars, vectors and matrices, respectively.
	$\boldsymbol{0}$ and $\boldsymbol{1}$ denote zero and one array of appropriate size, respectively. $\mathbb{I}^{x \times y}$, $\mathbb{R}_+^{x \times y}$, and $\mathbb{C}^{x \times y}$ denote the unit, real nonnegative, and complex spaces of dimension $x \times y$, respectively.
	$j$ denotes the imaginary unit.
	% $\arg(\cdot)$, $\mathrm{rank}(\cdot)$, $\mathrm{tr}(\cdot)$, $\mathrm{diag}(\cdot)$ and $\mathrm{diag}^{-1}(\cdot)$ denote the argument, rank, trace, a square matrix with input vector on main diagonal, and a vector retrieving main diagonal of input square matrix, respectively.
	$\mathrm{diag}(\cdot)$ returns a square matrix with the input vector on its main diagonal and zeros elsewhere.
	$(\cdot)^*$, $(\cdot)^T$, $(\cdot)^H$, $\lvert{\cdot}\rvert$, and $\lVert{\cdot}\rVert$ denote the conjugate, transpose, conjugate transpose, absolute value, and Euclidean norm operators, respectively.
	$(\cdot)^{(r)}$ and $(\cdot)^{\star}$ denote the $r$-th iterated and terminal solutions, respectively.
	The distribution of a \gls{cscg} random variable with zero mean and variance $\sigma^2$ is denoted by $\mathcal{CN}(0,\sigma^2)$, and $\sim$ means ``distributed as''.
\end{section}

@@

	\begin{subsection}{BackCom/SR: Backscatter Modulation}
		\gls{bc} and \gls{sr} commonly employ antenna-based scatterers as passive information sources.
		By \emph{frequently switching} between the available states, each backscatter node encodes self message over scattered signal and introduces backscatter uncertainty at the receiver.
		For $M$-ary \gls{qam}, reflection coefficient $\Gamma_m$ maps to the corresponding \emph{complex constellation point $c_m$} by
		% \cite{Thomas2012a}
		\begin{equation}
			\Gamma_m = \alpha \frac{c_m}{\max_{m'} \lvert c_{m'} \rvert},
			\label{eq:backscatter_modulation}
		\end{equation}
		where $\alpha \in \mathbb{I}$ is the amplitude scattering ratio at the direction of interest.
	\end{subsection}

	\begin{subsection}{RIS: Channel Reconfiguration}
		\gls{ris} typically involves a planar antenna array or metasurface of numerous metamaterial units as a passive channel reconfigurator.
		By \emph{adaptively choosing} the reflection pattern, each reflecting element alters the phase of scattered signal for constructive/destructive superposition at the receiver.
		For each unit with $M$ candidate states, reflection coefficient $\Gamma_m$ maps to the corresponding \emph{phase shift $\theta_m$} by
		% \cite{Wu2018}
		\begin{equation}
			\Gamma_m = \beta_m \exp(j \theta_m),
			\label{eq:passive_beamforming}
		\end{equation}
		where $\beta_m \in \mathbb{I}$ is the overall amplitude scattering ratio of state $m$.
	\end{subsection}

@@

\begin{section}{UniScatter}
	Using shared spectrum, energy and infrastructures, UniScatter is a novel passive communication/assistance protocol that generalizes \gls{bc}, \gls{sr} and \gls{ris} in a \emph{physically} and \emph{probabilistically} distributed manner.
	In particular, passive UniScatter nodes located at different places ride over an active legacy transmission in a flexible and mutualistic manner, while practical UniScatter receiver cooperatively decodes both primary and backscatter links under the benefits of backscatter modulation and passive beamforming.
	% TODO compare BC, SR and IRS
	\begin{subsection}{UniScatter Nodes}
		Based on link priority and \gls{csi}, UniScatter nodes adapt the \emph{state input probability distribution} of antenna-based scatterers to unify backscatter modulation and passive beamforming.
		The scattering sources of \gls{bc}/\gls{sr} and reflecting elements of \gls{ris} can be viewed as its extreme cases, where the input distribution boils down to equiprobable and degenerate, respectively.
		\begin{figure}[!t]
			\centering
			\subfloat[Input Distribution]{
				\resizebox{0.8\columnwidth}{!}{
					\input{assets/illustration/input_distribution.tex}
				}
				\label{fi:input_distribution}
			}
			\\
			\subfloat[Time Block]{
				\resizebox{0.8\columnwidth}{!}{
					\input{assets/illustration/time_block.tex}
				}
				\label{fi:time_block}
			}
			\caption{
				Input probability distribution and time block structure of \gls{bc}, \gls{sr}, \gls{ris}, and UniScatter.
				$T_s$ and $T_c$ respectively denote the primary and backscatter symbol period.
				Within channel coherence time, UniScatter nodes semi-randomly select reflection state for each backscatter symbol block under the guidance of input probability distribution.
			}
			\label{fi:uniscatter_distribution}
		\end{figure}
		As illustrated in Fig. \ref{fi:uniscatter_distribution}, instead of always using fully random or deterministic reflection pattern over time, each UniScatter node can \emph{flexibly adjust} its input distribution to balance information encoding and channel reconfiguration.
		That is, it semi-randomly select the reflection state per backscatter symbol block under the guidance of input probability $P(\Gamma_m)$ for state $m$.
		Such an adaptive backscatter channel coding coincides with \gls{ris} when primary link is absolutely prioritized, and outperforms the uncoded equiprobable inputs of \gls{bc}/\gls{sr} when backscatter link is absolutely prioritized.
		Besides, the physically distributed characteristics of UniScatter nodes and legacy users avoids the development optimization and allows uniformly good performance for both links.

@@

		Denote the \gls{ap}-user direct channel as $\boldsymbol{h}_{\mathrm{D}}^H \in \mathbb{C}^{1 \times Q}$, the \gls{ap}-node $k \in \mathcal{K} \triangleq \{1,\ldots,K\}$ forward channel as $\boldsymbol{h}_{\mathrm{F},k}^H \in \mathbb{C}^{1 \times Q}$, the node $k$-user backward channel as $h_{\mathrm{B},k}$, and the cascaded channel via tag $k$ as $\boldsymbol{h}_{\mathrm{C},k}^H \triangleq h_{\mathrm{B},k} \boldsymbol{h}_{\mathrm{F},k}^H \in \mathbb{C}^{1 \times Q}$.
		Let $x_{\mathcal{K}} \triangleq (x_1,\ldots,x_K)$ be the backscatter symbol tuple of all UniScatter nodes.
		Without loss of generality, we consider one specific backscatter block (i.e., $N$ primary blocks) in the following context.
		Due to double modulation, the equivalent primary channel is a function of \emph{coded} backscatter symbols
			\label{eq:equivalent_channel}
				\boldsymbol{h}_{\mathrm{E}}^H(x_{\mathcal{K}})
				 & \triangleq \boldsymbol{h}_{\mathrm{D}}^H + \sum_{k \in \mathcal{K}} \alpha_k \boldsymbol{h}_{\mathrm{C},k}^H x_k   \\
				 & = \boldsymbol{h}_{\mathrm{D}}^H + \boldsymbol{x}^H \mathrm{diag}(\boldsymbol{\alpha}) \boldsymbol{H}_{\mathrm{C}},
		where $\alpha_k \in \mathbb{I}$ is the amplitude scattering ratio of UniScatter node $k$, $x_k \in \mathcal{X} \triangleq \{c_1,\ldots,c_M\}$ is the coded backscatter symbol of node $k$, $\boldsymbol{\alpha} \triangleq [\alpha_1,\ldots,\alpha_K]^T \in \mathbb{I}^{K \times 1}$, $\boldsymbol{x} \triangleq [x_1,\ldots,x_K]^H \in \mathcal{X}^{K \times 1}$, and $\boldsymbol{H}_{\mathrm{C}} \triangleq [\boldsymbol{h}_{\mathrm{C},1},\ldots,\boldsymbol{h}_{\mathrm{C},K}]^H \in \mathbb{C}^{K \times Q}$. The signal received by the user at primary block $n \in \mathcal{N} \triangleq \{1,\ldots,N\}$ is

@@

			y[n] = \boldsymbol{h}_{\mathrm{E}}^H(x_{\mathcal{K}}) \boldsymbol{w} s[n] + v[n],
			\label{eq:receive_signal}
		where $s \sim \mathcal{CN}(0,1)$ is the primary symbol, $v \sim \mathcal{CN}(0,\sigma_v^2)$ is the \gls{awgn}, and $\boldsymbol{w} \in \mathbb{C}^{Q \times 1}$ is the active beamforming vector subject to average transmit power constraint $\lVert \boldsymbol{w} \rVert^2 \le P$.
		Denote $m_k \in \mathcal{M} \triangleq \{1,\ldots,M\}$ as the reflection state index of node $k$, and $m_{\mathcal{K}} \triangleq (m_1,\ldots,m_K)$ as the state index tuple of all nodes.
		Conditioned on backscatter index tuple $m_{\mathcal{K}}$, the received signal \eqref{eq:receive_signal} follows \gls{cscg} distribution $y[n] \sim \mathcal{CN}(0,\sigma_{m_{\mathcal{K}}}^2)$, where
			\sigma_{m_{\mathcal{K}}}^2 = \lvert \boldsymbol{h}_{\mathrm{E}}^H(x_{m_{\mathcal{K}}}) \boldsymbol{w} \rvert^2 + \sigma_v^2

@@

		is the received variance and $x_{m_\mathcal{K}}$ is the backscatter symbol tuple associated with $m_{\mathcal{K}}$.
		For node $k$, let $x_{m_k}$ be the backscatter symbol associated with index $m_k$.%

@@

			Please note $x_k$ and $x_{\mathcal{K}}$ are random variable and tuple, while $x_{m_k}$ and $x_{m_{\mathcal{K}}}$ are their instances of index $m_k$ and $m_{\mathcal{K}}$.

@@

		Besides, we define the total received energy per backscatter block as $z=\sum_{n=1}^N \lvert y[n] \rvert^2$.
		Since $z$ is the sum of $N$ \gls{iid} exponential variables, its \gls{pdf} conditioned on $m_{\mathcal{K}}$ follows Gamma distribution

@@

@@

		As illustrated in Fig. \ref{fi:energy_distribution}, the UniScatter receiver divides the received energy space into disjoint decision regions associated with those hypotheses.
		For example, if the total received energy during a backscatter block falls within region $\mathcal{R}_{m_{\mathcal{K}}}$, then the detector output would be $x_{m_{\mathcal{K}}}$.
		\begin{remark}
			Interestingly, the capacity-achieving decision design for \gls{dmtc} remains under-investigated, and some attempts were made for single source with binary inputs \cite{Qian2019b,Nguyen2021b}.
			For non-binary inputs with arbitrary distribution, the optimal decision region for each letter can be non-convex (i.e., with non-adjacent partitions) and the optimal number of thresholds is still unknown \cite{Nguyen2018,Nguyen2021}.
		\end{remark}
		Like most existing literatures, we limit the scope of this paper to convex decision regions and consider sequential decision thresholds design therein.
		% In particular, we define a general bijective mapping from backscatter index tuples to ordered indexes $f : m_{\mathcal{K}} \to \mathcal{L} \triangleq \{1,\ldots,L \triangleq M^K\}$, and use both notation interchangably in the following context.
		% In particular, we define a general bijective mapping from backscatter index tuples to ordered indexes $\mathcal{M}^K \to \mathcal{L} \triangleq \{1,\ldots,L \triangleq M^K\}$, $m_{\mathcal{K}} \mapsto l$ and use both notation interchangably in the following context.
		For the ease of notations, we define a general bijective function transforming backscatter index tuple $m_{\mathcal{K}}$ to integer $l \in \mathcal{L} \triangleq \{1,\ldots,L \triangleq M^K\}$, and use both notation interchangeably in the following context.
		As such, the convex decision region of backscatter index tuple $m_{\mathcal{K}}$ or equivalent letter $l$ can be written as
			\mathcal{R}_{m_{\mathcal{K}}} \triangleq [t_{l-1},t_l), \quad t_{l-1} \le t_l.
		Once the decision threshold vector $\boldsymbol{t} \triangleq [t_0,\ldots,t_L]^T \in \mathbb{R}_+^{(L+1) \times 1}$ is determined, we can formulate a \gls{dmtmac} with transition probability from input $x_{m_{\mathcal{K}}}$ to output $\hat{x}_{m_{\mathcal{K}}'}$

@@

			\label{eq:dmtmac}

@@

		is the regularized incomplete Gamma function that substitutes the \gls{dmtmac} transition probability \eqref{eq:dmtmac}.
